### 2017-03-02 Meetings/KoM02March17/kom_020317_crossexperiment.tex
added the suggestion from Jesn and te cross-machine slide

\documentclass[10pt, compress]{beamer}
\usetheme[conference=KoM,venue=Remote, date=03/13/2017, titleprogressbar, logo=RFX-logo]{Eurof}
\usepackage{listings,amsmath,multimedia, amssymb}
\usepackage{../beamerclass/tangocolors}
\usepackage{../beamerclass/rfxcolor}
% for drawing
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds}
\usepackage{../beamerclass/onimage}
\usepackage[export]{adjustbox}
\usepackage{bm}
% for font
\usepackage[absolute,overlay]{textpos}
  \setlength{\TPHorizModule}{1mm}
  \setlength{\TPVertModule}{1mm}

\usepackage[style=nature,citestyle=authoryear-comp,defernumbers=true,maxnames=2,firstinits=true,
uniquename=init,backend=bibtex8,arxiv=abs,mcite]{biblatex}
\bibliography{biblio}
\renewcommand*{\bibfont}{\footnotesize}
\renewcommand*{\citesetup}{\footnotesize}
\usepackage[export]{adjustbox}
\makeatother
\mode<presentation>
\makeatletter
% add a macro that saves its argument
\newcommand{\footlineextra}[1]{\gdef\insertfootlineextra{#1}}
\newbox\footlineextrabox
% for reducing font on a single slide
\newcommand\Fontvi{\fontsize{8}{7.2}\selectfont}
\title{Topic 21: AUG experiment KoM}
\date{03 March 2017}
\author[N.Vianello]{N. Vianello}
\begin{document}
\tikzstyle{every picture}+=[remember picture]
\maketitle

\begin{frame}{Cross-machine comparison}
\vspace{-1cm}
  \begin{itemize}
    \item Brief summary of proposals done by N.Vianello and
      F. Militello who suggest the repetition of similar experiments
      in the 3 devices to provide appropriate comparison
    \item<2|only@2> \textbf{L-Mode}
      \begin{itemize}
        \item Current scan both at fixed and varied q$_{95}$
        \item Equivalent density ramp in favorable/unfavorable
          $B\times \nabla B$ direction
        \item Probe at a fixed position for long time statistics
        \item Parallel flow and modification accordingly to
          favorable/unfavorable direction
        \end{itemize}
      \item<3|only@3> \textbf{H-Mode}
      \begin{itemize}
        \item H-Mode with density ramps/seeding proposed by all the proponents
        \item Similar density ramps/seeded experiments performed in
          LSN/USN/DN configuration and monitor of filaments at the
          target and at the midplane
        \item QH-mode or Type-III ELMy regime
        \end{itemize}
\end{itemize}
\end{frame}
\end{document}