### 2017-05-11 Meetings/TCVrelated/kom_150517_vianello.tex
started to prepare slides for TCV KOM Meeting

\documentclass[10pt, compress]{beamer}
\usetheme[conference=TCV-Topic21 KoM,venue=Lausanne, date=15/05/2017, titleprogressbar, logo=RFX-logo]{Eurof}
\usepackage{listings,amsmath,multimedia, amssymb}
\usepackage{../beamerclass/tangocolors}
\usepackage{../beamerclass/rfxcolor}
% for drawing
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds}
\usepackage{../beamerclass/onimage}
\usepackage[export]{adjustbox}
\usepackage{bm}
% for font
\usepackage[absolute,overlay]{textpos}
  \setlength{\TPHorizModule}{1mm}
  \setlength{\TPVertModule}{1mm}

\usepackage[style=nature,citestyle=authoryear-comp,defernumbers=true,maxnames=2,firstinits=true,
uniquename=init,backend=bibtex8,arxiv=abs,mcite]{biblatex}
\bibliography{biblio}
\renewcommand*{\bibfont}{\footnotesize}
\renewcommand*{\citesetup}{\footnotesize}
\usepackage[export]{adjustbox}
\makeatother
\mode<presentation>
\makeatletter
% add a macro that saves its argument
\newcommand{\footlineextra}[1]{\gdef\insertfootlineextra{#1}}
\newbox\footlineextrabox
% for reducing font on a single slide
\newcommand\Fontvi{\fontsize{8}{7.2}\selectfont}
\title{Topic 21: TCV experiment KoM}
\date{15 March 2017}
\author[N. Vianello,  V. Naulin]{N . Vianello and V. Naulin}
\begin{document}
\tikzstyle{every picture}+=[remember picture]
\maketitle

\begin{frame}{TCV experiments: boundary condition}
\vspace{-1cm}
\Fontvi
\only<1>{
  \begin{itemize}
  \item 2017 \textbf{objectives} listed after the General Planning Meeting
    \begin{enumerate}
      \item Provide cross-machine \alert{L-Mode} shoulder dependence on
        current both at constant Bt and at constant q$_{95}$
      \item Establish robust scenario for density shoulder profile in
        H-mode and establish dependence on fuelling/neutral
        profiles/divertor condition
      \item Study the role of ELM regimes, neutral compression, and
        particle density in filamentary transport and related shoulder
        formation.
      \item Identify the contribution of collisionality and
        seeding on filamentary transport and related shoulder
        formation.
      \item Determine the effect of filaments and shoulder
        formation on target heat loads in different Hmode plasmas.
    \end{enumerate}
  \item We have a total number of \textbf{\# 23 Shots} split into two
    operational window. Calendar week 24 (12.06-16.06) and Calendar
    week 43 (23.10-27.10)
\end{itemize}
}
\only<2>{
  \begin{itemize}
    \item 15 people listed in the team for TCV experiment
      \begin{description}
        \item[EPFL:] H. De Oliveira, R. Maurizio, B. Labit,
          C. Tsui, K. Verhaegh, H. Reimerdes, C. Theiler
        \item[DTU:] J.J. Rasmussen and V. Naulin
        \item[RFX:] N. Vianello, M. Spolaore, M. Agostini
        \item[OEAW:] B. Schneider, S. Costea, C. Ionita
        \item[CCFE:] F. Militello
        \item[JSI:] J. Kovacic
      \end{description}
    \end{itemize}
  }
\end{frame}
\begin{frame}{Results obtained during TCV15-2.2-3 Experiment}

\end{frame}
% \only<2>{
% We will conclude the \textcolor{ta3chameleon}{L-Mode}
% part in Week 15 where we also would like to include part of the \textcolor{ta3scarletred}{H-Mode}
% scenario development. This would give us an additional week between
% the experiments to address possible issues in particular for the
% H-mode part\\
% \textbf{Week 15} \\
% We choose are reference shot \# 30269 (I$_p$ = 0.8 MA, B$_{\phi}$ = 2.5T, q$_{95}\approx 4.5$)
%   \begin{enumerate}
%     \item \textcolor{ta3chameleon}{Shot at I$_p$=0.6 MA
%       keeping the same toroidal field $B_{\phi}$ = 2.5T as the reference adjust the fueling
%       rate (\emph{Current scan with modification of $q_{95}$})}
%   \item \textcolor{ta3chameleon}{Shot at I$_p$ = 0.6MA
%       reduce the toroidal field B$_{\phi}\approx $2T in order to match q$_{95}\approx 4.5$, 
%       (\emph{Current scan at fixed $q_{95}$}). This would accommodate
%       the diagnostic request of D. Aguiam}
%     \item \textcolor{ta3chameleon}{Density ramp with I$_p$ = 1MA and
%         B$_{\phi}$ = 2.5T}
%     \item \textcolor{ta3chameleon}{Density ramp with I$_{p}$ = 1MA,
%         increasing the toroidal field in order to match q$_{95}\approx
%         4.5$}
%     \end{enumerate}
%   }
%   \only<3>{
%   \textbf{Week 15} \\
%     For the \textcolor{ta3scarletred}{H-Mode} scenario development we start from the best shot found in
%     2016 (\# 33478) and increase the heating power
%     \begin{enumerate}
%       \item \textcolor{ta3scarletred}{Start from shot \# 33478 but
%           with increased heating power (6MW). Adjust fueling rate from
%         reference by increasing by a factor of 30 \%. 1 Plunge of
%         probe head at the end of the discharge still in a safe
%         position and IR monitoring}
%     \item \textcolor{ta3scarletred}{Repeat \# 1 eventually adjusting
%         the fueling rate. Start the N seeding in feedforward starting
%         fro the level found in reference}
%     \item \textcolor{ta3scarletred}{Trade off between \#1 and \#2
%         Fueling/Seeding. Additional plunge of probe at the end of the
%         discharge}
%       \item \textcolor{ta3skyblue}{This scenario would allow the
%           exploitation of particle accelleration (McClements) physics as piggy-back}
%     \end{enumerate}
%   }
%   \only<4>{
%   \textbf{Week 17: This strongly depends on the achievement of H-Mode
%     scenario obtained in Week 15 } \\
%     \begin{enumerate}
%     \item \textcolor{ta3scarletred}{Repeat best H-Mode shot found in Week 15 1st Radial position of probe}
%     \item \textcolor{ta3scarletred}{Repeat \#1, different probe position}
%     \item \textcolor{ta3scarletred}{Repeat \#1, different probe
%         position}
%     \item \textcolor{ta3scarletred}{Repeat best H-Mode shot found in
%         week 15 and reduced the cryopumps}
%     \item \textcolor{ta3scarletred}{Repeat best H-Mode shot found in
%         week 15 and puff from midplane}
%     \item \textcolor{ta3scarletred}{Contingency}
%     \item \textcolor{ta3scarletred}{Contingency}
%     \end{enumerate}
%   }
%   \only<5>{
%     Among the contingency we propose the following 4 possibilities to
%     be discussed
%     \begin{enumerate}
%     \item \textcolor{ta3chameleon}{Reversing B$_t$ direction and repeat one identical shot (e.g \# 30269) to investigate
%    the role of SOL flows in SOL shoulder formation and filamentary
%    transport}
%     \item \textcolor{ta3chameleon}{DN discharge with similar density ramps as in reference. Possibly the two X-point should
%    sit on the same flux surface}
%    \item \textcolor{ta3scarletred}{Attempt a scenario similar to Topic-06 which will be performed later in time. See for example
%    shot \# 29816 (Presented by T. Eich in the GPM) which is at even
%    higher power (8 MW) or \# 25740
%    which is actually in DN (see proposal from J. Vicente). If we choose for this we could actually compare with the priority 3 of L-Mode
%    contingency}
%  \item \textcolor{ta3scarletred}{Reverse B$_{t}$ operation. In this case the L-H treshold is different and we might end
%      by careful adjusting the power into I-Mode scenario}
%    \item \textcolor{ta3skyblue}{Working at 2T in H-Mode would require
%        additional development}
%    \item \textcolor{ta3skyblue}{Given that the upper divertor is less
%        diagnosed then the lower one we prefer eventually to operate
%        with reverse B$_{\phi}$}

% \end{enumerate}
%   }
% \end{frame}
% \begin{frame}{Required diagnostic and analysis}
%   \begin{itemize}
%     \item[$\boxtimes$] Midplane Manipulator
%     \item[$\square$] Li-Beam. \emph{Are fluctuations and profiles
%         available simultaneously}
%     \item[$\square$] RFA \#2
%     \item[$\square$] Divertor probes
%     \item[$\boxtimes$] Neutral profiles
%     \item[$\square$] Infrared for probe head monitoring. \emph{Are Target
%       infrared measurements available/useful?}
%     \item[$\square$] GPI
%     \item[$\boxtimes$] Reflectometer. \emph{The operation at 2T can be
%       obtained during the q$_{95}$ scan}
%     \item[$\boxtimes$] Fast probes on the limiter
%     \item[$\square$] Bolometer/AXUV in the divertor region
%   \end{itemize}
% \end{frame}

% \begin{frame}{Work to be accomplished before the experiment}
%   \begin{enumerate}
%     \item Check the shape modification during the current/q$_{95}$
%       scan
%     \item Probe conditioning?
%     \item Check the status of the diagnostics including GPI (issue
%       regarding the puffing)
%     \item Methodology ($\Lambda_{div}$ computation and profile)
%       $\lambda_n$ filaments properties etc.  
%     \item Code preparation for analysis and
%       visualization. \emph{GITHUB repository?}
%     \item Optimization of the effort: please provide us a more
%       detailed plan of your analysis in order to limit superposition
%       or work duplication
%   \end{enumerate}
% \end{frame}
\end{document}