### 2017-05-30 Meetings/TCVrelated/revised_shotlist.tex
Added a short presentation for revised TCV shotlist

\documentclass[10pt, compress]{beamer}
\usetheme[conference=TCV-Topic21,venue=Lausanne, date=30/05/2017, titleprogressbar, logo=RFX-logo]{Eurof}
\usepackage{listings,amsmath,multimedia, amssymb}
\usepackage{../beamerclass/tangocolors}
\usepackage{../beamerclass/rfxcolor}
% for drawing
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds}
\usepackage{../beamerclass/onimage}
\usepackage[export]{adjustbox}
\usepackage{bm}
% for font
\usepackage[absolute,overlay]{textpos}
  \setlength{\TPHorizModule}{1mm}
  \setlength{\TPVertModule}{1mm}

\usepackage[style=nature,citestyle=authoryear-comp,defernumbers=true,maxnames=2,firstinits=true,
uniquename=init,backend=bibtex8,arxiv=abs,mcite]{biblatex}
\bibliography{biblio}
\renewcommand*{\bibfont}{\footnotesize}
\renewcommand*{\citesetup}{\footnotesize}
\usepackage[export]{adjustbox}
\makeatother
\mode<presentation>
\makeatletter
% add a macro that saves its argument
\newcommand{\footlineextra}[1]{\gdef\insertfootlineextra{#1}}
\newbox\footlineextrabox
% for reducing font on a single slide
\newcommand\Fontvi{\fontsize{8}{7.2}\selectfont}
\title{Topic 21: TCV experimental plan for Week 24}
\date{30 March 2017}
\author[N. Vianello,  V. Naulin]{N . Vianello and V. Naulin}
\begin{document}
\tikzstyle{every picture}+=[remember picture]
\maketitle

\begin{frame}{Taking advantage of already performed experiment}
\vspace{-1cm}
\Fontvi
  \begin{columns}
    \begin{column}{0.6\textwidth}
      \includegraphics[width=\textwidth]{../../Experiments/TCV/analysis/pdfbox/CurrentScanTopic25}
    \end{column}
    \begin{column}{0.4\textwidth}
      \begin{itemize}
        \item Topic 25 already performed a current scan at
          constant B$_{\phi}$ but as far as I checked we don't have
          fast reciprocating probe data for all the scan \alert{To be
            confirmed by Cedric}
        \item Furthermore the density was not optimized for lower
          current, as it start to high
      \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Proposed density ramp}
\Fontvi
  \begin{columns}
    \begin{column}{0.6\textwidth}
      \includegraphics[width=\textwidth]{../../Experiments/TCV/analysis/pdfbox/ProposedDensityRamp}
    \end{column}
    \begin{column}{0.4\textwidth}
      \begin{itemize}
        \item I propose to perform with the same density ramp at all
          the current as shown in dashed line. Less demanding for the
          higher current case we could end up into disruption for the
          lower current case
      \end{itemize}
    \end{column}
  \end{columns}
  \end{frame}

  \begin{frame}{Proposed experimental plan}
\Fontvi
    \alert{For the first week of operation we propose to perform only
  L-Mode shot and we have accommodated the low collisionality request
  from cedric}
\begin{enumerate}
\item Shape from 57088, I$_p$ = 245 kA,  Reverse B$_t$,
    density ramp from Line Average Density = 3.8e19 @ 0.5 s to 11e19 @ 1.6s,  Bt = 1.4T. Plunge @ 0.65, 1.52
\item  Repeat \# 1 with I$_p$=330 kA Bt=1.4T, same density ramp, same timing for plunges
\item  Repeat \# 1 with I$_p$=180 kA, Bt=1.4T, same density ramp, same timing for plunges
\item  Repeat \# 1 with q95=2.44 as \# 2, adjust Bt consequently (Bt = 1.02T)
\item  Repeat \# 3 with q95=2.44 as \# 2, adjust Bt consequently (Bt=0.8T)
\item  Shape and current from \# 1. Stop puffing once the divertor is formed to get low collisionality case
\item  Repeat \# 6 with density feedback controlled at value 50 \%
  higher then \# 6. 
\item  Repeat density ramp of Shot \# 2 in DN configuration 
\item  Repeat density ramp of Shot \# 3 in DN configuration 
\item Repeat \# 1 in forward field
\item Repeat \# 3 in forward field
\end{enumerate}
\end{frame}
\end{document}